Automatic Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-08 paper/paper_new.tex
Tweaked Fig. 10 placement.
Edited abstract.

%abstract can be shortened a good bit.
%can offer an alternative manana..
\begin{abstract}
Metadata is a key data source, particularly for researchers seeking to apply machine learning (ML) to the vast collections of digitized specimens. Unfortunately, the available metadata is often sparse and, at times, erroneous. Additionally, it is prohibitively expensive to address these limitations through traditional, manual means.
This paper reports on research that applies machine-driven approaches to
analyzing digitized fish images and extracting various important
features from them. The digitized fish specimens are being analyzed as part of the Biology Guided Neural Networks (BGNN) initiative, which is developing a novel class of artificial neural networks using phylogenies and anatomy ontologies. Automatically generated metadata is crucial for identifying
the high-quality images needed for the neural network's predictive analytics.
%predictive analytics?
Methods that  combine ML and image informatics techniques allow us to rapidly enrich the existing metadata associated with each of 
7,244 images from the Illinois Natural History Survey (INHS). The sample
images all have 22 manually generated metadata properties associated with them, which
aid our study.  Results show we can accurately generate many key metadata properties relevant to the
BGNN project such as: fish quantity and orientation,  image scaling based on
ruler identification and measurement, and general image quality metrics
(e.g.\ brightness and contrast). Results also show we can also accurately generate bounding boxes and segmentation masks for fish, which are needed for subsequent machine learning analysis.
%to confirm list of six properties.
%x, y, z = any other key results. jg will add from results section, unless someone else does first. 
The automatic process out-performs humans in terms of time and labor cost,
as well as accuracy, and provides a novel solution for leveraging digitized
specimens in ML. This research has implications for harnessing the tens of thousands of digitized specimens stored in open-access repositories found wold-wide.
%world wide [*across the globe] or we may not even need this part as open access is global access

@@

deep learning
% modified AlexNet \cite{AlexNet2012}

@@

\subsection{Scale and Length}
Image scale and fish lengths were calculated for \(7,179\) of the images.
For the remaining 65 images, either the fish, the ``2'' and/or the ``3'' on the ruler were not detected.
Image scale ($\frac{\mathrm{pixels}}{\mathrm{cm}}$) and fish length were
measured, using ImageJ~\cite{imagejCite}, in the same 100  test images.
% 50 of these images \todo{The same 50 as before?} were measured by hand to get \(\frac{\mathrm{pixels}}{\mathrm{cm}}\) and fish length in centimeters with ImageJ~\cite{imagejCite}.
In this subset of images the average error for the scale calculation
was \(0.89\%\), and the average error for the fish length calculation was \(5.55\%\).
Scale calculations using the ``2'' and ``3'' method are nearly identical to those calculated by hand between the tick marks on the ruler. When the tail of the fish is accurately masked and the specimen is fairly straight, the length calculation is highly accurate as well. An example of such a result can be seen in Figure~\ref{fig:scale_len}, for which the difference between the hand measured length and the automatically calculated length was only \(0.6\%\) (\(8.88\) cm vs \(8.82\) cm). Thus, the primary means of lowering the error of the length calculation is to improve tail masking accuracy.
\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/54172}
  \includegraphics[width=0.49\linewidth]{images/54172_mask}
  \caption{A fish for which the masking and measurement process was highly accurate.}
  \label{fig:scale_len}
\end{figure}

>>>>

%abstract can be shortened a good bit.
%can offer an alternative manana..
\begin{abstract}
Metadata is a key data source, particularly for researchers seeking to apply machine learning (ML) to the vast collections of digitized specimens. Unfortunately, the available metadata is often sparse and, at times, erroneous. Additionally, it is prohibitively expensive to address these limitations through traditional, manual means.
This paper reports on research that applies machine-driven approaches to
analyzing digitized fish images and extracting various important
features from them. The digitized fish specimens are being analyzed as part of the Biology Guided Neural Networks (BGNN) initiative, which is developing a novel class of artificial neural networks using phylogenies and anatomy ontologies. Automatically generated metadata is crucial for identifying
the high-quality images needed for the neural network's predictive analytics.
%predictive analytics?
Methods that  combine ML and image informatics techniques allow us to rapidly enrich the existing metadata associated with the
7,244 images from the Illinois Natural History Survey (INHS) used in our
study. The sample
images also have 22 manually generated metadata properties, which
can be used for comparison.  Results show we can accurately generate many key metadata properties relevant to the
BGNN project such as: fish quantity and orientation,  image scaling based on
ruler identification and measurement, and general image quality metrics
(e.g.\ brightness and contrast). Results also show we can also accurately generate bounding boxes and segmentation masks for fish, which are needed for subsequent machine learning analysis.
%to confirm list of six properties.
%x, y, z = any other key results. jg will add from results section, unless someone else does first. 
The automatic process out-performs humans in terms of time and labor cost,
as well as accuracy, and provides a novel solution for leveraging digitized
specimens in ML. This research demonstrates the ability of computational
methods to enhance the digital library services associated with
the tens of thousands of digitized specimens stored in open-access
repositories world-wide.
%world wide [*across the globe] or we may not even need this part as open access is global access

@@

% deep learning
modified AlexNet \cite{AlexNet2012}

@@

\begin{figure}[b]
  \centering
  \includegraphics[width=0.49\linewidth]{images/54172}
  \includegraphics[width=0.49\linewidth]{images/54172_mask}
  \caption{A fish for which the masking and measurement process was highly accurate.}
  \label{fig:scale_len}
\end{figure}

\subsection{Scale and Length}
Image scale and fish lengths were calculated for \(7,179\) of the images.
For the remaining 65 images, either the fish, the ``2'' and/or the ``3'' on the ruler were not detected.
Image scale ($\frac{\mathrm{pixels}}{\mathrm{cm}}$) and fish length were
measured, using ImageJ~\cite{imagejCite}, in the same 100  test images.
% 50 of these images \todo{The same 50 as before?} were measured by hand to get \(\frac{\mathrm{pixels}}{\mathrm{cm}}\) and fish length in centimeters with ImageJ~\cite{imagejCite}.
In this subset of images the average error for the scale calculation
was \(0.89\%\), and the average error for the fish length calculation was \(5.55\%\).
Scale calculations using the ``2'' and ``3'' method are nearly identical to those calculated by hand between the tick marks on the ruler. When the tail of the fish is accurately masked and the specimen is fairly straight, the length calculation is highly accurate as well. An example of such a result can be seen in Figure~\ref{fig:scale_len}, for which the difference between the hand measured length and the automatically calculated length was only \(0.6\%\) (\(8.88\) cm vs \(8.82\) cm). Thus, the primary means of lowering the error of the length calculation is to improve tail masking accuracy.
